### 2021-04-26 input/proposal.tex

%% to better adapt positioning (I expect 2.5 pages
%% intro+related work)

% \vfill \TODO{spacing for intro and related work: 2.5 pages}
% \newpage

@@

%% In this section, we propose to dynamically adapt the scopes of \NAMEB
%% using \NAMEB itself for the sake of efficiency: a \NAMEB message may
%% modify the state of a \process, consequently changing predicate
%% evaluations over deliveries.  At any time, a \process decides whether
%% or not it becomes the center (or \emph{source}) of a scope (or
%% \emph{partition}). \Processes must eventually rally the partition
%% corresponding to their closest source despite concurrent operations,
%% and dynamic changes in the network.

%% \noindent In that regard, we introduce \NAME (stands for
%% \underline{A}daptive \underline{S}coped broad\underline{cast}), a
%% wait-free reactive protocol for dynamic logical partitioning that
%% enables adaptive scoped broadcast in dynamic distributed
%% systems. \NAME's operation guarantees that \processes eventually reach
%% consistent partitioning by efficiently propagating their state
%% changes. \NAME's overhead actually depends on its operations and
%% current partitions in the system. When the system becomes quiescent,
%% \processes eventually converge to their respective partition and do not
%% require further communication afterward.

%% \noindent This section starts by introducing consistent partitioning
%% and its implementation when a \process can only become a new source to
%% create a new partition, i.e., it cannot revoke its self-appointed
%% status of source. Then, it extends to include the latter behavior: it
%% introduces the properties that guarantee consistent partitioning in
%% such context, and it details our implementation called \NAME along
%% with its complexity analysis.

To assign and maintain \processes to their best partition according to
replica creations/removals as well as dynamic infrastructure changes,
we designed and implemented \NAME (\NAME stands for
\underline{A}daptive \underline{S}coped broad\underline{cast}).  \NAME
relies on a primitive that allows a \process to broadcast a message
within a limited scope.  This primitive is used in a first algorithm
to deliver consistent partitioning each time \process creates a new
replica within the system. This algorithm is then extended to include
replica removals as well as dynamic changes of the infrastructure.
This section discussed the different properties that guarantee
consistent partitioning in such a context and details the \NAME
implementation along with its complexity analysis.

@@

% \TODO{dynamic graph: superscript with $t$?}
In the edge infrastructure we consider, nodes are heterogeneous and
interconnected by communication links. \Processes involved in the
management of the data sets may crash but are not byzantine.  Finally,
\processes can reliably communicate through asynchronous message
passing to other known \processes called neighbors.
A more formal definition of our system is given below:

@@

We define the Scoped broadcast (\NAMEB) as a communication primitive that propagates a
message around its broadcaster within an application-dependant scope :

@@

This definition encompasses more specific definitions of related
work~\cite{hsiao2005scoped, lue2006scoped, wang2015prodiluvian}.  It
also highlights that epidemic propagation and scoped broadcast have
well-aligned preoccupations. More precisely, it underlines the transitive relevance of
messages, thus \processes can stop forwarding messages as soon as the
corresponding predicate becomes unverified.
%For instance, a \process
%from Paris could scoped broadcast messages to all \processes in
%Paris. This requires \processes to store and maintain their city
%location in their local state. \Processes stop delivering and
%forwarding their received messages when they come from a different
%city.  In another instance, a \process from Paris could scoped
%broadcast messages to all \processes in Paris plus neighboring
%cities. This requires \processes to overload forwarded messages the
%first time they reach another city. The predicate checks if messages
%already reached two distinct cities before delivery. Similarly to
%uniform reliable broadcast, scoped broadcast implementations expose
%different trade-offs on space, time, and communication.

\NAMEB is used to change the state of each \process dynamically, depending on all partitions that exist in the system.

@@

At any time, a \process can decide to become a source, hence creating
a new partition in the system by executing an \texttt{Add}
operation. This partition includes at least its source plus
neighboring \processes that estimate they are closer to this source
than any other one. Such distance (or \emph{weight}) is
application-dependant: in the context of maintaining distributed
indexes, weights would be link latency.

@@

\begin{definition}[\underline{C}onsistent \underline{p}artitioning (CP)]
  Let $W_{xy} = W_{yx}$ be the positive symmetric \underline{w}eight
  between $x$ and $y$, $\Pi_{xz}$ be the shortest \underline{p}ath
  from $x$ to $z$ the weight of which $|\Pi_{xz}|$ is lower than any
  other path weight, the only consistent partitioning $\mathcal{P}$ is
  a set of partitions $P_{s\in S}$ such that each \process belongs to
  a logical partition comprising its closest source: $\forall p \in
  P_{s_1}, \nexists P_{s_2}$ such that $|\Pi_{s_1p}| > |\Pi_{s_2p}|$.
\end{definition}

Unfortunately, \processes do not share a common global knowledge of
the network state. For \processes to reach consistent partitioning
together, a \process becoming a source must notify every \process that
is closer to it than any other source. Since epidemic dissemination
and scoped broadcast have well-aligned preoccupations, we assume
implementations relying on forwarding of messages from
neighbor-to-neighbor.

@@

At any time, a source can revoke its self-appointed status of source
by executing a \texttt{Del} operation, hence deleting its partition
from the system. All \processes that belong to this partition must
eventually choose another partition to belong to. In
Figure~\ref{fig:del}, two partitions exist initially: $P_a$ and $P_d$
that respectively include $\{a\}$, and $\{b, c, d\}$. In
Figure~\ref{fig:delA}, \Process~$a$ deletes its partition. It notifies
all neighboring \processes -- here only \Process~$b$ -- that may
belong to its partition using \NAMEB. Upon receipt, \Process~$b$
discards the message $\delta_a$, for the latter does not target the
former's partition. \Process~$b$ sends its own knowledge of best
partition $\alpha_d^3$ that may prove to be the best for
\Process~$a$. Eventually, every \processes belong to the same
partition $P_d$. In this scenario, they reach consistent partitioning.

@@

Adding an edge between two \processes may only improve the shortest
path of one of these \processes. Therefore, it triggers a competition
between the two \processes only. If one or the other \process
improves, the propagation of $\alpha$ messages operates normally.
Removing an edge between two \processes may invalidate the shortest
path of one of involved \processes plus downstream \processes. As a
side effect, removing an edge may also impair the propagation of a
partition delete. To implement this behavior, \NAME uses its
implementation of $\delta?$ messages. This prove costly but
enables \NAME even in dynamic networks subject to physical
partitioning.

>>>>

@@

To assign and maintain \processes to their best partition according to
replica creations and removals, as well as dynamic infrastructure
changes, we designed and implemented \NAME.  \NAME stands for
\underline{A}daptive \underline{S}coped broad\underline{cast}. It
relies on a primitive that allows a \process to broadcast a message
within a limited scope. We first use this primitive to guarantee
consistent partitioning when a \process can only create a new replica
within the system. We highlight the issue when a process can also
destroy a replica, and provide a second algorithm that handles replica
removals as well as dynamic changes of the infrastructure.  This
section describes the communication primitive called scoped broadcast,
then discusses the properties that guarantee consistent partitioning
in our context, and finally details our implementation \NAME and its
complexity.

@@

% \TODO{dynamic graph: superscript with $t$?}
In this paper, we consider edge infrastructures with heterogeneous
\nodes interconnected by communication links. \Processes involved in
the management of content may crash but are not byzantine.  Finally,
\processes can reliably communicate through asynchronous message
passing to other known \processes called neighbors.  % A more formal
%definition of our system is given below:

@@

We define scoped broadcast as a communication primitive that
propagates a message around its broadcaster within an
application-dependant scope:

@@

This definition encompasses more specific definitions of related
work~\cite{hsiao2005scoped, lue2006scoped, wang2015prodiluvian}.  It
also highlights that epidemic propagation and scoped broadcast have
well-aligned preoccupations. More precisely, it underlines the transitive relevance of
messages, thus \processes can stop forwarding messages as soon as the
corresponding predicate becomes unverified.
%For instance, a \process
%from Paris could scoped broadcast messages to all \processes in
%Paris. This requires \processes to store and maintain their city
%location in their local state. \Processes stop delivering and
%forwarding their received messages when they come from a different
%city.  In another instance, a \process from Paris could scoped
%broadcast messages to all \processes in Paris plus neighboring
%cities. This requires \processes to overload forwarded messages the
%first time they reach another city. The predicate checks if messages
%already reached two distinct cities before delivery. Similarly to
%uniform reliable broadcast, scoped broadcast implementations expose
%different trade-offs on space, time, and communication.
%
We use \NAMEB to dynamically and efficiently modify the state of each
\process depending on all partitions that exist in the system.

@@

At any time, a \process can decide to become a \emph{source}, hence
creating a new partition in the system by executing an \texttt{Add}
operation. This partition includes at least its source plus
neighboring \processes that estimate they are closer to this source
than any other one. Such distance (or \emph{weight}) is
application-dependant: in the context of maintaining distributed
indexes, weights would be link latency.

@@

\begin{definition}[\underline{C}onsistent \underline{p}artitioning (CP)]
  Let $W_{xy} = W_{yx}$ be the positive symmetric \underline{w}eight
  between $x$ and $y$, $\Pi_{xz}$ be the shortest \underline{p}ath
  from $x$ to $z$ the weight of which $|\Pi_{xz}|$ is lower than any
  other path weight, the only consistent partitioning $\mathcal{P}$ is
  a set of partitions $P_{s\in S}$ such that each \process belongs to
  a logical partition comprising its closest source: $\forall p \in
  P_{s_1}: \nexists P_{s_2}$ such that $|\Pi_{s_1p}| > |\Pi_{s_2p}|$.
\end{definition}

Unfortunately, \processes do not share a common global knowledge of
the network state. For \processes to reach consistent partitioning
together, a \Process $s$ \underline{a}dding a partition must send a
notification $\alpha_s$ to every \process that is closer to it than
any other source. Since epidemic dissemination and scoped broadcast
have well-aligned preoccupations, we assume implementations relying on
message forwarding from neighbor-to-neighbor.

@@

At any time, a source can revoke its self-appointed status of source
by executing a \texttt{Del} operation, hence deleting its partition
from the system. All \processes that belong to this partition must
eventually choose another partition to belong to. In
Figure~\ref{fig:del}, two partitions exist initially: $P_a$ and $P_d$
that respectively include $\{a\}$, and $\{b, c, d\}$. In
Figure~\ref{fig:delA}, \Process~$a$ deletes its partition. It notifies
all neighboring \processes~--~here only \Process~$b$~-- that may
belong to its partition using \NAMEB. Upon receipt, \Process~$b$
discards the \underline{d}elete notification $\delta_a$ since
$\delta_a$ does not target the former's partition $P_d$. \Process~$b$
sends its own best partition $\alpha_d^3$ that may be the best for
\Process~$a$. Eventually, every \process belongs to Partition
$P_d$. In this scenario, they reach consistent partitioning.

@@

Adding an edge between two \processes may only improve the shortest
path of one of these \processes. Therefore, it triggers a competition
between the two \processes only. If one or the other \process
improves, the propagation of $\alpha$ messages operates normally.

\noindent Removing an edge between two \processes may invalidate the
shortest path of one of involved \processes plus downstream
\processes. As a side effect, removing an edge may also impair the
propagation of a partition delete. To implement this behavior, \NAME
uses its implementation of $\delta?$ messages. This prove costly but
enables \NAME even in dynamic networks subject to physical
partitioning.
